---
title: "Dynamic Documents in R"
author: "Tham, Wei Yang"
date: "March 2, 2016"
output: 
  ioslides_presentation:
    incremental: true
---

## R and R Studio {.build}
Getting started 

- Download R and RStudio (first exit off the Google highway!)
- a change done locally in my Mac
- a change done on server
- Download R and RStudio (first exit off the Google highway!)
- Learn with the `swirl` package (<http://swirlstats.com/>)

Why R?

- Free
- `ggplot2` (graphics) and `dplyr` (data wrangling) packages
- Dynamic documents - easy to embed code into documents
  + My examples are for R Markdown, which can be converted into pdf 
  + This can all be done in LaTex directly
  
## Embedding code - 1 {.build}

- Embedding code in documents is easy
- Just type:

<pre><code>```{r chunk_name}
x <- 1 + 1
y <- 6
x/y
```</code></pre>

```{r, ex1}
x <- 1 + 1
y <- 6
x/y
```

## Embedding code - 2 {.build}

- Code can be embedded inline too
```{r ex2, message = F, warning = F}
# load the "nycflights13" package, which contains flights data
library(nycflights13)
library(dplyr)
flights %>% select(year, month, arr_delay, dep_delay, carrier) %>% 
head(., 3)
```

## Embedding code - 3 {.build}

- I want to say: The flights dataset has N observations and K variables
```{r ex2a}
dim(flights)
```

- The `flights` dataset has `` `r '\x60r dim(flights)[1]\x60'` ``  observations and `` `r '\x60r dim(flights)[2]\x60'` `` variables
- The `flights` dataset has `r dim(flights)[1]` observations and `r dim(flights)[2]` variables
- No more cutting and pasting!

## Plots {.build}
```{r fig1, fig.align = 'center'}
library(ggplot2)
qplot(carat, price, data = diamonds %>% filter(cut == "Ideal"))
```

## Regression tables - 1 
```{r ex3a, echo = F, message = F}
library("AER") # Applied Econometrics with R
library("stargazer")
library(dplyr)
library(nycflights13)

daily <- flights %>%
  filter(origin == "EWR") %>%
  group_by(year, month, day) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE))

daily_weather <- weather %>%
  filter(origin == "EWR") %>%
  group_by(year, month, day) %>%
  summarise(temp   = mean(temp, na.rm = TRUE),
            wind   = mean(wind_speed, na.rm = TRUE),
            precip = sum(precip, na.rm = TRUE))

# Merge flights with weather data frames
both <- inner_join(daily, daily_weather) %>% 
  data.frame()  # Temporary fix

# Create an indicator for quarter
both$quarter <- cut(both$month, breaks = c(0, 3, 6, 9, 12), 
                                labels = c("1", "2", "3", "4"))

# Create a vector of class logical
both$hot <- as.logical(both$temp > 85)
```

Lots of options for generating reg tables e.g. stargazer (get it?)
```{r ex3b, eval = F, results = 'asis', align = 'center'}
output  <- lm(delay ~ temp + wind + precip, data = both)
output2 <- lm(delay ~ temp + wind + precip + quarter, data = both)
stargazer(output, output2, type = "html", keep.stat = c("n", "adj.rsq"), 
omit = "quarter", style = "aer")
```

## Regression tables - AER {.flexbox .vcenter}
```{r ex3c, echo = F, results = 'asis'}
output  <- lm(delay ~ temp + wind + precip, data = both)
output2 <- lm(delay ~ temp + wind + precip + quarter, data = both)
stargazer(output, output2, type = "html", keep.stat = c("n", "adj.rsq"),  
omit = "quarter", style = "aer")
```

## Regression tables - QJE {.flexbox .vcenter}
```{r ex3d, echo = F, results = 'asis'}
output  <- lm(delay ~ temp + wind + precip, data = both)
output2 <- lm(delay ~ temp + wind + precip + quarter, data = both)
stargazer(output, output2, type = "html", keep.stat = c("n", "adj.rsq"),  
omit = "quarter", style = "qje")
```

## References

- [Jake Russ's stargazer cheatsheet](http://jakeruss.com/cheatsheets/stargazer.html)
- Hlavac, Marek (2015). stargazer: Well-Formatted Regression and Summary Statistics Tables. R
package version 5.2. <<http://CRAN.R-project.org/package=stargazer>>